### 2018-03-05 doc/ms.tex
More writing

\documentclass[a4paper, 12pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs} % For nicer tables
\usepackage{color}
\usepackage[left=2cm, right=2cm, bottom=3cm, top=2cm]{geometry}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{microtype}
\usepackage{natbib}
\usepackage{transparent}

\renewcommand{\arraystretch}{1.2} % More space between table rows

\title{Bayesian Inference from Time Series: Information Loss and the Whittle Likelihood}
\author{Brendon J. Brewer}
\date{2018}

\begin{document}
\maketitle

\abstract{\noindent I investigate the information content of a time series
          dataset for inferring a parameter of interest. }

% Need this after the abstract
\setlength{\parindent}{0pt}
\setlength{\parskip}{1em}

\section{A Damped, Stochastically-Excited Oscillator}
Consider the following ordinary differential equation, which describes
a damped and excited oscillator:
\begin{align}
\frac{d^2y}{dt^2} + \frac{\omega_0}{Q}\frac{dy}{dt}
        + \omega_0^2\frac{dy}{dt} &= \epsilon(t).\label{eqn:ode}
\end{align}
The function $y(t)$ is the resulting signal, and the parameters $\omega_0$
are the angular frequency (that the signal would have if
the driving force $\epsilon(t)$ were zero) and $Q$ is the quality factor.
If the driving force $\epsilon(t)$ is only defined probabilistically (i.e.,
it is specified to be white noise) then Equation~\ref{eqn:ode} becomes a
stochastic differential equation whose solution is a probability distribution
for $y(t)$.

We assume that the probability distribution for $\epsilon(t)$ is indeed white
noise. In this case, the probability distribution for $y(t)$ given the
parameters is a gaussian process with mean function of zero and
covariance function given by
\begin{align}
C(\tau) &= A^2\left[\cos(\eta\omega_0\tau)
                + \frac{1}{2\eta Q}\sin(\eta\omega_0\tau)\right]
\end{align}
where $\eta = \sqrt{1 - 1/(4Q^2)}$, $\tau$ is a time separation,
and $Q > 1/2$ (i.e., we assume the oscillation is not overdamped).
A more general expression which allows for $Q \leq 1/2$
is given by \citet{anderson} and \citet{celerite}.
In the limit of large $Q$, the covariance function is approximately
the product of an oscillation and an exponential decay:
\begin{align}
C(\tau) &= A\exp\left(-\frac{\omega_0\tau}{2Q}\right)\cos(\omega_0\tau)
\end{align}
which was the form assumed by \citep{brewer_stello}.

\begin{table}[!ht]
\centering
\begin{tabular}{@{}l@{\hspace{3em}}l@{\hspace{2em}}l@{}}
\toprule
$N$         &       Likelihood      &  Mut. Inf. (nats) \\
\hline
20          &       Exact           & \\
20          &       Whittle         & $1.491 \pm 0.055$\\ 
50          &       Exact           & \\
50          &       Whittle         & $1.957 \pm 0.055$ \\
100         &       Exact           & $2.605 \pm 0.053$ \\
100         &       Whittle         & $2.052 \pm 0.055$ \\
200         &       Exact           &   \\
200         &       Whittle         &   \\
\bottomrule
\end{tabular}
\caption{\label{tab:oscillator_results}}
\end{table}

\begin{table}[!ht]
\centering
\begin{tabular}{@{}l@{\hspace{3em}}l@{}}
\toprule
$N$         &       $\Delta$ Mut. Inf. (nats) \\
\hline
50          & \\
100         &       $0.553 \pm 0.056$ \\
200         & \\
\bottomrule
\end{tabular}
\caption{The difference in mutual information between the exact and
Whittle likelihoods. This quantifies the size of the advantage of the exact
likelihood.
The error bars are smaller than what you might expect from
taking differences in Table~\ref{tab:oscillator_results}, due to
the common stream of datasets.\label{tab:oscillator_differences}}
\end{table}

\begin{thebibliography}{999}
\bibitem[Anderson et al.(1990)]{anderson}
Anderson, Edwin R., Thomas L. Duvall Jr, and Stuart M. Jefferies. ``Modeling of solar oscillation power spectra.'' The Astrophysical Journal 364 (1990): 699-705.

\bibitem[Brewer and Stello(2009)]{brewer_stello}
Brewer, Brendon J., and Dennis Stello. ``Gaussian process modelling of asteroseismic data.'' Monthly Notices of the Royal Astronomical Society 395, no. 4 (2009): 2226-2233.

\bibitem[Foreman-Mackey et al.(2017)]{celerite}
Foreman-Mackey, D., Agol, E., Ambikasaran, S. and Angus, R., 2017. Fast and scalable Gaussian process modeling with applications to astronomical time series. The Astronomical Journal, 154(6), p.220.
\end{thebibliography}

\end{document}